\chapter{Further Observations and Conclusion}
\label{chap:observations}

% ─────────────────────────────────────────────────────────────────────────────
\section{General Electric's Reverse Split Adjustment}
\label{sec:ge_jump}

During initial analysis, General Electric appeared to be the top-performing stock with a massive 940\% return, driven by a near-vertical price jump from $\sim$\$22 to $\sim$\$182 between July 30 and August 2, 2021. 

\textbf{Identifying the anomaly:} Unlike the 10 corrupted rows where only \texttt{high}/\texttt{close} were artificially inflated and immediately reverted the next day, GE's jump involved all four price columns shifting simultaneously and establishing a persistent new price level. Viewed in isolation, this price-only evidence is ambiguous: it could be a genuine (if extreme) market repricing or a corporate action. 

To break the ambiguity, trading volume was inspected. On the day of the price jump, GE's reported volume dropped by exactly a factor of 8, perfectly mirroring the 8x price increase. This simultaneous, inversely proportional shift in both price and volume is the classic signature of a 1-for-8 reverse stock split, confirming it as a structural change rather than a sudden change in company valuation.

\textbf{Important caveat:} While a reverse split is the most financially sound interpretation of this specific pattern, in a purely synthetic dataset, this could theoretically represent another data generation bug (e.g., an errant multiplier accidentally applied to both price and volume for all rows before August 2). However, given the exact inverse relationship and the permanence of the shift, treating it as a reverse split is the correct analytical decision.

\textbf{Why it was adjusted:} While a reverse split mathematically increases the per-share price, it proportionately reduces the number of shares an investor owns. It does not represent true organic performance or portfolio growth. To ensure the final ranking reflected true market performance, all of General Electric's historical prices prior to August 2 were multiplied by 8 (and volume divided by 8). This realigned the historical baseline, revealing GE's true organic return to be less than 30\%, allowing Lockheed Martin to rightfully enter the Top 5.

% ─────────────────────────────────────────────────────────────────────────────
\section{Signs of Synthetic Data Design}
\label{sec:synthetic}

Several characteristics of the dataset point clearly to its synthetic origin:

\begin{itemize}
    \item \textbf{Round-number returns:} several top performers have suspiciously exact returns, such as Netflix (180.00\%), Nokia (110.00\%), and NVIDIA (85.00\%). General Electric, before adjustment, also sat at exactly 940.00\%. Round percentages are extremely unlikely to occur naturally across multiple independent stocks in a single time window. This strongly suggests the dataset was designed with specific target returns for a subset of ``winner'' stocks.

    \item \textbf{Discrete price jumps as the mechanism:} the data generator appears to have implemented target returns via single-day price jumps rather than a smooth exponential curve. General Electric's raw +694\% daily move (which we adjusted) was a clear example of this. This is a convenient shortcut in synthetic data construction but would be essentially impossible in real equity markets.

    \item \textbf{Systematic data corruption at a fixed ratio:} the 10 corrupted rows all shared an identical multiplier of approximately 11 across unrelated stocks. Real data errors at this scale would be unlikely to produce such a consistent signature. This pattern is more consistent with a data generation bug that applied a specific transformation to a randomly selected subset of rows.

    \item \textbf{Organic-looking noise elsewhere:} outside the top performers and the corrupted rows, the remaining stocks show realistic day-to-day volatility with no round-number patterns, which suggests a genuine effort was made to make the bulk of the data behave like real market data.
\end{itemize}

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}

The five best-performing stocks from June 1 to October 13, 2021 were Netflix (180.0\%), Adobe (141.49\%), Nokia (110.0\%), NVIDIA (85.0\%), and Lockheed Martin (69.85\%). These results are supported by a rigorous, multi-step data quality process that identified and corrected 10 systematically inflated price rows, adjusted historical prices for a 1-for-8 reverse stock split (General Electric), interpolated 23 missing records per-ticker, and resolved a corrupted CSV header before any calculations were performed.

The analysis also demonstrated several principles that matter in real-world data work: the importance of distinguishing data entry errors from legitimate corporate actions (and adjusting for them mathematically), the value of stating ambiguous terms explicitly before computing them (30 trading days vs.\ 30 calendar days), and the risk of reusing a filtered dataset for a calculation that requires a wider data range (the volume lookback extending before the analysis window).

The full code is available in \texttt{hardik\_src/solution.ipynb} for reference.
